---
title: "OmniVerse: Multi-Omics RNA Analysis Shiny Application"
author: "Safa Rafique"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{OmniVerse}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

**OmniVerse** is a Shiny application for high-throughput genomic analysis of bulk RNA-seq and microarray data (e.g. from GEO). It integrates with Bioconductor packages (GEOquery, Biobase, limma, DESeq2, edgeR, clusterProfiler) in a single workflow: download, QC, normalization, batch correction, differential expression, WGCNA, pathway enrichment, PPI, machine learning, and optional immune deconvolution.

# Installation

Install OmniVerse from Bioconductor (use `BiocManager`):

```{r install, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
  install.packages("BiocManager")
BiocManager::install("OmniVerse")
```

# Running the application

After installation, launch the app with `runOmniVerse()`. The app will open in your default browser.

```{r run, eval=FALSE}
library(OmniVerse)
runOmniVerse(launch.browser = TRUE)
```

To run on a specific port or allow external access:

```{r run-port, eval=FALSE}
runOmniVerse(port = 3838, host = "0.0.0.0")
```

# Workflow overview

The app guides you through:

1. **Download** – Fetch RNA-seq or microarray data from GEO (e.g. GSE IDs).
2. **Groups** – Define sample groups (e.g. Normal vs Disease).
3. **QC** – Quality control and filtering.
4. **Normalize** – Normalization (e.g. TPM, log2 for microarrays).
5. **Batch correction** – Optional ComBat or other batch adjustment.
6. **Differential expression** – DESeq2, limma, or edgeR.
7. **Results** – Tables, volcano plots, and export.
8. **WGCNA** – Co-expression modules (optional).
9. **Pathways** – Enrichment (e.g. KEGG, GO) and common genes.
10. **PPI** – Protein–protein interaction and hub genes.
11. **Machine learning** – Feature selection, ROC, nomogram, validation.
12. **Immune** – Optional immune cell deconvolution.

# More information

- **R help:** In R, run `?runOmniVerse` for the function help and references.
- **In the app:** Each step has an **"About this step"** or info box (click the header to expand) with purpose, methods, and tips. Use the **Guided Tour** button in the app header to highlight the main workflow; you can use it to capture a full-page screenshot for your own documentation.
- **Session info:** The section below shows the R session and package versions used when this vignette was built. Use it to reproduce the environment or report issues.

# Session information

This chunk runs when the vignette is built and records the R version and loaded packages.

```{r session}
sessionInfo()
```
