---
title: "OmniVerse: Multi-Omics RNA Analysis Workflow"
author: "Safa Rafique"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    toc: true
    toc_float: true
vignette: >
  %\VignetteIndexEntry{OmniVerse: Multi-Omics RNA Analysis Workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

**OmniVerse** is a Shiny application for high-throughput genomic analysis of bulk RNA-seq and microarray data. It integrates with the Bioconductor ecosystem (GEOquery, Biobase, limma, DESeq2, edgeR, clusterProfiler, and others) to provide a single workflow: from GEO download and QC, through normalization and differential expression, to pathway enrichment, WGCNA, protein–protein interaction, machine learning, and optional immune deconvolution.

This vignette describes the workflow and how to run the app.

# Installation

Install OmniVerse from source (e.g. after cloning the repository or unpacking the release):

```{r install, eval=FALSE}
install.packages(".", repos = NULL, type = "source")
# Or with devtools:
# devtools::install()
```

Load the package:

```{r load}
library(OmniVerse)
packageVersion("OmniVerse")
```

# Running the application

Launch the Shiny app. It opens in your default browser and listens on port 3838 by default.

```{r run-app, eval=FALSE}
runOmniVerse()
runOmniVerse(launch.browser = TRUE, port = 3838, host = "127.0.0.1")
```

# Workflow overview

The app guides you through:

1. **Download** — Fetch GEO series (GSE) and platform (GPL) data; optional CEL file download for microarrays.
2. **QC** — Sample and gene-level QC, PCA, correlation.
3. **Normalization** — RMA (oligo/affy) for microarrays; TMM/DESeq2-style for RNA-seq.
4. **Batch correction** — Optional ComBat (sva) when multiple batches are present.
5. **Differential expression** — limma (microarray and RNA-seq) or DESeq2/edgeR for RNA-seq.
6. **Downstream** — WGCNA, pathway enrichment (clusterProfiler), PPI (STRINGdb), ML (e.g. random forest, XGBoost), optional immune deconvolution, validation.

Data are kept in memory as Bioconductor-friendly structures (e.g. `ExpressionSet`, count matrices, and design tables) and can be saved/loaded as RDS for later use.

# Interoperability with Bioconductor

OmniVerse uses standard Bioconductor classes and accessors:

- **Microarray:** `Biobase::ExpressionSet`, `exprs()`, `fData()`, `GEOquery::getGEO()` and `Table()` for GPL annotations.
- **RNA-seq:** `edgeR::DGEList`, `DESeq2::DESeqDataSetFromMatrix()` and related DESeq2/edgeR workflows.

This allows pipelines and custom scripts to consume the same objects produced or used by the app.

# Session info

```{r session-info}
sessionInfo()
```
